---
title: "Actual FI"
author: "Ross Finnegan"
date:  '`r format(Sys.time(), "%H:%M:%S on %d %B, %Y")`'
output: 
  html_document:
    number_sections: TRUE
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
options(width = 128)
knitr::opts_chunk$set(echo = TRUE)
```

# Setup

```{r results='hide'}
library("asreml")
source("qgm_helpers_t2.R")
```


```{r results='hide'}
packageVersion("asreml")
```
# Reading phenotypic data and pedigree

```{r}
actAFI <- read.csv(file="data_for_genan.csv",header=TRUE,na.strings=c("NA"))
head(actAFI)
tail(actAFI)
```
```{r}
ped <- read.csv(file= "ped_gen0to6.csv", header= TRUE, na.strings=c("NA"))
head(ped)
tail(ped)
```

# Preparing for the analysisn - Data cleaning
```{r}
actAFI$icowi <- actAFI$icow

actAFI$sample_date_formatted <- as.Date(actAFI$sample_date, format =  "%d/%m/%Y %H:%M")
actAFI$YEAR <- format(actAFI$sample_date_formatted, format="%Y")
actAFI$MONTH <- format(actAFI$sample_date_formatted, format="%M")

```


## Declaring factors
We need to define factors in the analysis. 
```{r}
actAFI$GENETIC_GROUP <- factor(actAFI$GENETIC_GROUP)
actAFI$FEED_TYPE <- factor(actAFI$FEED_TYPE)
actAFI$LACT_NO <- factor(actAFI$LACT_NO)
actAFI$icow <- factor(actAFI$icow) 
actAFI$icowi <- factor(actAFI$icowi)
actAFI$YEAR <- factor(actAFI$YEAR)
actAFI$MONTH <- factor(actAFI$MONTH)
#actAFI$idam <- factor(actAFI$idam)
```

## Distributional assumptions for random effect `id`

The distributional assumptions for `id` will account for the relationships among all the individuals and will be distributed $MVN(0,V_a \mathbf{A})$ where $V_a$ is the additive genetic variance and $\mathbf{A}$ is Wright's Numerator Relationship Matrix. 
 
```{r}
ainv <- ainverse(ped[,1:3]) 
```

```{r}
setwd("C:/Users/rossf/Downloads")
fullset <- read.csv(file="Fullset2.csv",header=TRUE,na.strings=c("NA"))

```
```{r}
qgm_ped_summary(ped[,1:3])
```




# Fitting the Individual (Animal) Model

## Running the model

Ready to fit the Individual (Animal) Model in `asreml()` and estimate a variance component.

 * The form has 4 elements: `fixed=`, `random=`, `residual=` and `data=`. 
 * For the Individual (Animal) Model we need to specify that the random term is `id` and the covariance of the breeding values, i.e. the numerator relationship matrix, is specified using `vm(id,ainv)`.
 * The residual term has been kept as `idv(units)`, to be explicit that the covariance model of the residuals is the 'identity' matrix i.e. independent and identically distributed.

```{r}
## Basic Animal Model
actAFI1.asr <- asreml(fixed= weeklyave_FI ~ 1 + GENETIC_GROUP + FEED_TYPE + LACT_NO +  YEAR:MONTH, random= ~ vm(icow,ainv), residual= ~ idv(units), data=actAFI)

## Animal Model with environmental and maternal effect modelled
actAFI2.asr <- asreml(fixed= weeklyave_FI ~ 1 + GENETIC_GROUP + FEED_TYPE + LACT_NO +  YEAR:MONTH, random= ~ vm(icow,ainv) + idv(icowi) , residual= ~ idv(units), data=actAFI)
```

## ... and looking at the output summary

```{r}
qgm_summary(actAFI1.asr)
```
```{r}
qgm_summary(actAFI2.asr)
```

## Estimates of Heritability

Using `vpredict()` to estimate the phenotypic variance as the sum of the variance components i.e. $V_a + V_e$. 

 
```{r}
# phenotypic variance & heritability?
vpredict(actAFI1.asr, vp ~ V1+V2)
vpredict(actAFI1.asr, h ~ V1/ (V1+V2))


vpredict(actAFI2.asr, vp ~ V1+V2+V3)
vpredict(actAFI2.asr, h ~ V2/(V1+V2+V3))
```

# Looking at the EBVs

## Extract and merge EBVs

The estimates obtained from fitting an Individual (Animal) Model are the 'Expected Breeding Values' or 'Estimated Breeding Values', often denoted EBVs. 
The EBVs can be extracted from the ASReml output using the `qgm_random_effects()` function.

```{r}
ebv <- qgm_random_effects(actAFI.asr,"id")
head(ebv)
tail(ebv)
nrow(ebv)
```


```{r}
colnames(ebv)
colnames(ebv)[1] <- "ebv"
colnames(ebv)[2] <- "se"
colnames(ebv)[4] <- "id" # rename ebv$levels as ebv$id to match with ped data frame ready for merging
pedx <- merge(ped, ebv, by = "id")
head(pedx)
tail(pedx)
```


## Wald tests

```{r}
print("WALD TESTS")
print(as.data.frame(wald(actAFI.asr,denDF="numeric",ssType="conditional")$Wald)[,c(1,2,4,6)])
```
